Derogations and Democratic Backsliding: Exploring the Pandemic's Effects on Civic Spaces

### 2021-03-23 manuscript/manuscript.Rmd
Super neat way more efficient way of knitting

---
title: "Derogations and Democratic Backsliding: Exploring the Pandemic's Effects on Civic Spaces"
author:
- name: Suparna Chaudhry
  affiliation: Lewis & Clark College
  email: schaudhry@lclark.edu
  url: http://www.suparnachaudhry.com/
- name: Andrew Heiss
  affiliation: Georgia State University
  email: aheiss@gsu.edu
  url: https://www.andrewheiss.com/
date: March 21, 2021
published: This is an early draft of a working paper.
git-repo: https://github.com/andrewheiss/covid-derogations
abstract: >-
  In an effort to combat the COVID-19 pandemic, most governments have imposed restrictions on civic freedoms in the interest of public health. In many cases, these emergency measures have been temporary and proportional in terms of controling the pandemic. In many other instances, however, governments have used these restrictions to suppress opposition and more permanently restrict civic space. Systematically measuring the consequences of COVID restrictions, however, is a difficult task. We examine two possible quantitative measures of the relationship between of COVID restrictions and civil society space. First, we use the Variety of Democracy project's newly released Pandemic Violations of Democratic Standards Index to explore if and how civil society restrictions predict pandemic backsliding. Second, while many countries sign international human rights treaties that ostensibly bind states to respect rights, several treaties allow for emergency derogations from these obligations. We tabulate formal human rights treaty derogations due to the pandemic and explore whether these emergency measures led to lasting declines in associational and human rights. We hope that our exploration of these two measures will provide rich data-based descriptions of the relationship between COVID restrictions and civic space that will allow for more causal work in the future.
keywords: "Thing 1, Thing 2, Thing 3"
reference-section-title: References
bibliography: bibliography.bib
csl: pandoc/csl/chicago-author-date.csl
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.retina = 3,
                      tidy.opts = list(width.cutoff = 120),  # For code
                      options(width = 90),  # For output
                      fig.asp = 0.618, fig.width = 7, 
                      fig.align = "center", out.width = "85%",
                      echo = FALSE, warning = FALSE, message = FALSE)

options(dplyr.summarise.inform = FALSE,
        knitr.kable.NA = "")

>>>>

---
title: "Derogations and Democratic Backsliding: Exploring the Pandemic's Effects on Civic Spaces"
short-title: "Derogations and Democratic Backsliding"
author:
- name: Suparna Chaudhry
  affiliation: Lewis & Clark College
  email: schaudhry@lclark.edu
  url: http://www.suparnachaudhry.com/
- name: Andrew Heiss
  affiliation: Georgia State University
  email: aheiss@gsu.edu
  url: https://www.andrewheiss.com/
date: March 21, 2021
published: This is an early draft of a working paper.
git-repo: https://github.com/andrewheiss/covid-derogations
abstract: >-
  In an effort to combat the COVID-19 pandemic, most governments have imposed restrictions on civic freedoms in the interest of public health. In many cases, these emergency measures have been temporary and proportional in terms of controling the pandemic. In many other instances, however, governments have used these restrictions to suppress opposition and more permanently restrict civic space. Systematically measuring the consequences of COVID restrictions, however, is a difficult task. We examine two possible quantitative measures of the relationship between of COVID restrictions and civil society space. First, we use the Variety of Democracy project's newly released Pandemic Violations of Democratic Standards Index to explore if and how civil society restrictions predict pandemic backsliding. Second, while many countries sign international human rights treaties that ostensibly bind states to respect rights, several treaties allow for emergency derogations from these obligations. We tabulate formal human rights treaty derogations due to the pandemic and explore whether these emergency measures led to lasting declines in associational and human rights. We hope that our exploration of these two measures will provide rich data-based descriptions of the relationship between COVID restrictions and civic space that will allow for more causal work in the future.
keywords: "international human rights treaties, civil society, COVID-19, democracy"
reference-section-title: References
bibliography: bibliography.bib
csl: pandoc/csl/chicago-author-date.csl
title-page: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
# Figure out output format
is_docx <- knitr::pandoc_to("docx") | knitr::pandoc_to("odt")
is_latex <- knitr::pandoc_to("latex")
is_html <- knitr::pandoc_to("html")

# Word-specific things
table_format <- ifelse(is_docx, "huxtable", "kableExtra")  # Huxtable tables
conditional_dpi <- ifelse(is_docx, 300, 300)  # Higher DPI
conditional_align <- ifelse(is_docx, "default", "center")  # Word doesn't support align

# Knitr options
knitr::opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE,
  tidy.opts = list(width.cutoff = 120),  # Code width
  fig.retina = 3, dpi = conditional_dpi,
  fig.width = 7, fig.asp = 0.618,
  fig.align = conditional_align, out.width = "85%",
  fig.path = "output/figures/",
  cache.path = "output/_cache/",
  fig.process = function(x) {  # Remove "-1" from figure names
    x2 = sub('-\\d+([.][a-z]+)$', '\\1', x)
    if (file.rename(x, x2)) x2 else x
  }
)

# R options
options(
  width = 90,  # Output width
  dplyr.summarise.inform = FALSE,  # Turn off dplyr's summarize() auto messages
  knitr.kable.NA = "",  # Make NAs blank in kables
  kableExtra.latex.load_packages = FALSE,  # Don't add LaTeX preamble stuff
  modelsummary_default = table_format  # Set modelsummary backend
)
